Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-01-07 Manuscript/TEM_draft.tex
Outlined and drafted discussion

\subsection{Effect of missing data for fossil taxa ($M_{F}$)}
%When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa and leads to displacement of certain taxa. However, it is important to note that even though the number of displaced wildcard taxa increases (i.e. decrease of Normalised Triplets distance), the conservation of the clades (i.e. Normalised Robinson-Foulds distance) is still relatively good (mode=% value
%) when the proportion of missing data is high ($M_{F}$=75\%).

%Methods
%Decrease in clade conservation is smaller than for param ML in the Bayesian consensus tree for the clade conservation but higher than param ML for the displacement of wildcard taxa. This is due to the fact that the Bayesian consensus tree is a majority consensus rule and that when the fossils have less data, they kind of randomly branch all over the shop and therefore no majority consensus position (>50\% of the time) exists puting the fossil on a politomie at the base of a clade (clade conservation stays high but displacement of taxa increases). On the opposite in MaxLik, the topology is always dichotomous but the fossil can be missplaced in the wrong clade (with a low BS value) but is moving less. In other words, Bayesian consensus tree places the fossil with a higher confidence at a low taxonomic level (politomy) and the MaxLik places the fossil with a low confidence at a high taxomomic level (baldy supported dichotomies). We can argue that the first option (higher confidence at a low taxonomic level) is preferable (example the Darwinius debate).

\subsection{Effect of missing character ($M_{C}$)}
%When the overall number of morphological characters in the matrix ($M_{C}$) decreases

%Story for MC: MC is expected to have an effect two times bigger than ML or MF, e.g. removing 10\% of MF removes 5\% of data (because it works only on the fossil part) and removing 10\% of MC actually removes 10\% of that.
%Method
%However the effect of removing characters has an effect of topology of the same order of magnitude than for ML and MF (FigRes 1): For RF, Bayesian consensus is better than ML and for Tr, MaxLik is better than Baycon and have the same drop at 50\% as for MF (different than for ML).

%In this scenario, the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}). % NC: What's the relevance of that sentence?
%However, the pairwise Bhattacharyya Coefficients show that the area with the lowest probability of distributional overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}). %NC: I don't understand what you mean here

\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
%In our simulations, the overlap in data is greatly reduced when all three missing data parameters have a lot of missing data (e.g. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
%When combining the effect, we

>>>>

\subsection{Effect of missing data for fossil taxa ($M_{F}$)}
When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa and leads to displacement of certain taxa. However, it is important to note that even though the number of displaced wildcard taxa increases (i.e. decrease of Normalised Triplets distance), the conservation of the clades (i.e. Normalised Robinson-Foulds distance) is still relatively good (mode=% value
%) when the proportion of missing data is high ($M_{F}$=75\%).

%Methods
Decrease in clade conservation is smaller than for param ML in the Bayesian consensus tree for the clade conservation but higher than param ML for the displacement of wildcard taxa. This is due to the fact that the Bayesian consensus tree is a majority consensus rule and that when the fossils have less data, they kind of randomly branch all over the shop and therefore no majority consensus position (>50\% of the time) exists puting the fossil on a politomie at the base of a clade (clade conservation stays high but displacement of taxa increases). On the opposite in MaxLik, the topology is always dichotomous but the fossil can be missplaced in the wrong clade (with a low BS value) but is moving less. In other words, Bayesian consensus tree places the fossil with a higher confidence at a low taxonomic level (politomy) and the MaxLik places the fossil with a low confidence at a high taxomomic level (baldy supported dichotomies). We can argue that the first option (higher confidence at a low taxonomic level) is preferable (example the Darwinius debate).

\subsection{Effect of missing character ($M_{C}$)}
When the overall number of morphological characters in the matrix ($M_{C}$) decreases
Story for MC: MC is expected to have an effect two times bigger than ML or MF, e.g. removing 10\% of MF removes 5\% of data (because it works only on the fossil part) and removing 10\% of MC actually removes 10\% of that.
%Method
However the effect of removing characters has an effect of topology of the same order of magnitude than for ML and MF (FigRes 1): For RF, Bayesian consensus is better than ML and for Tr, MaxLik is better than Baycon and have the same drop at 50\% as for MF (different than for ML). 
In this scenario, the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}). % NC: What's the relevance of that sentence?
However, the pairwise Bhattacharyya Coefficients show that the area with the lowest probability of distributional overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}). %NC: I don't understand what you mean here

\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
In our simulations, the overlap in data is greatly reduced when all three missing data parameters have a lot of missing data (e.g. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
When combining the effect, we
